---
title: "RCTs PSET 2"
author: "Matthew Zhao"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    keep_tex: yes
header-includes:
  \usepackage{dcolumn}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,lubridate,stargazer,xtable,kableExtra)
```

## 1 Randomization and Balance Checks

### 1.1

Brune et al. (2021) asks what is the impact of offering a deferred savings program to workers on saving/investment outcomes and what are the mechanisms which affect take-up and use. 

### 1.2

```{r}
# clear environment
rm(list = ls())

# set seed
set.seed(2090820)

# import data
df_complete <- read.csv('data/brune_et_al_data.csv')

# complete randomization
n <- df_complete %>% nrow()

# n draws from uniform(0,1), essentially assigning a random number to each worker
# when combined with rank
df_complete$rand <- runif(n, 0, 1)
df_complete$rank <- rank(df_complete$rand)

# assigns top half of random rank to control, bottom to treatment
df_complete$treat <- ifelse(df_complete$rank <= (n / 2), 1, 0)

table(df_complete$treat)
```

### 1.3

```{r results=F}
complete_summary <- df_complete %>%
  group_by(treat) %>%
  summarise(female_w = sum(female_w),
            d_married_w = sum(d_married_w),
            worker_age_w = mean(worker_age_w),
            plucker_w = sum(plucker_w),
            sav_all_tot_tot_w = mean(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat,names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat, names_prefix = 'treat_', 
              values_from = value)

complete_stats <- df_complete %>%
  select(treat, female_w, d_married_w, 
         worker_age_w, plucker_w, sav_all_tot_tot_w) %>%
  pivot_longer(cols=!treat,
               names_to='name',values_to='value') %>%
  group_by(name) %>%
  summarise(
            tstat = t.test(value ~ treat)$statistic,
            pval = t.test(value ~ treat)$p.value,
            .groups='keep')

complete_bal <- left_join(complete_summary,
                          complete_stats,
                          by='name')
complete_bal_tab <- xtable(complete_bal)
align(complete_bal_tab) <- xalign(complete_bal_tab)
display(complete_bal_tab) <- xdisplay(complete_bal_tab)
print(complete_bal_tab, include.rownames=FALSE)
```

\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
name & treat\_0 & treat\_1 & tstat & pval \\ 
  \hline
female\_w & 159.00 & 142.00 & 1.21 & 0.23 \\ 
  d\_married\_w & 301.00 & 308.00 & -0.52 & 0.61 \\ 
  worker\_age\_w & 39.56 & 39.47 & 0.13 & 0.90 \\ 
  plucker\_w & 341.00 & 326.00 & 1.20 & 0.23 \\ 
  sav\_all\_tot\_tot\_w & 31968.41 & 32160.47 & -0.06 & 0.96 \\ 
   \hline
\end{tabular}
\end{table} 


### 1.4

```{r}
# stratified randomization

# clear environment
rm(list = ls())

# set seed
set.seed(2090820)

# import data
df_strat <- read.csv('data/brune_et_al_data.csv')

n <- df_strat %>% nrow()
df_strat$rand <- runif(n, 0, 1)

df_strat <- df_strat %>%
  group_by(division) %>%
  mutate(treat_strat = ifelse(rand <= quantile(rand,c(0.33)), 0, 1))

df_strat %>% 
  count(division, treat_strat) %>% 
  pivot_wider(names_from = treat_strat, 
              names_prefix = 'treat_',
              values_from = n)
```

### 1.5

```{r results=F}
strat_summary <- df_strat %>%
  group_by(treat) %>%
  summarise(female_w = sum(female_w),
            d_married_w = sum(d_married_w),
            worker_age_w = mean(worker_age_w),
            plucker_w = sum(plucker_w),
            sav_all_tot_tot_w = mean(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat,names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat, names_prefix = 'treat_', 
              values_from = value)

strat_stats <- df_strat %>%
  select(treat, female_w, d_married_w, 
         worker_age_w, plucker_w, sav_all_tot_tot_w) %>%
  pivot_longer(cols=!treat,
               names_to='name',values_to='value') %>%
  group_by(name) %>%
  summarise(tstat = t.test(value ~ treat)$statistic,
            pval = t.test(value ~ treat)$p.value,
            .groups='keep')

strat_bal <- left_join(strat_summary,
                          strat_stats,
                          by='name')
strat_bal_tab <- xtable(strat_bal)
align(strat_bal_tab) <- xalign(strat_bal_tab)
display(strat_bal_tab) <- xdisplay(strat_bal_tab)
print(strat_bal_tab, include.rownames=FALSE)
```

