---
title: "RCTs PSET 2"
author: "Matthew Zhao"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    keep_tex: yes
header-includes:
  \usepackage{dcolumn}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,lubridate,stargazer,xtable,kableExtra)
```

## 1 Randomization and Balance Checks

### 1.1

Brune et al. (2021) asks what is the impact of offering a deferred savings program to workers on saving/investment outcomes and what are the mechanisms which affect take-up and use. 

### 1.2

```{r}
# clear environment
rm(list = ls())

# set seed
set.seed(2090820)

# import data
df_complete <- read.csv('data/brune_et_al_data.csv')

# complete randomization
n <- df_complete %>% nrow()

# n draws from uniform(0,1), essentially assigning a random number to each worker
# when combined with rank
df_complete$rand <- runif(n, 0, 1)
df_complete$rank <- rank(df_complete$rand)

# assigns top half of random rank to control, bottom to treatment
df_complete$treat <- ifelse(df_complete$rank <= (n / 2), 1, 0)

table(df_complete$treat)
```

### 1.3

```{r results=F}
# table of means with each row as a variable and T/C as columns
complete_means <- df_complete %>%
  group_by(treat) %>%
  summarise(female_w = mean(female_w),
            d_married_w = mean(d_married_w),
            worker_age_w = mean(worker_age_w),
            plucker_w = mean(plucker_w),
            sav_all_tot_tot_w = mean(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat,names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat, names_prefix = 'treat_', 
              values_from = value)

# table of standard deviations 
complete_sds <- df_complete %>%
  group_by(treat) %>%
  summarise(female_w = sd(female_w),
            d_married_w = sd(d_married_w),
            worker_age_w = sd(worker_age_w),
            plucker_w = sd(plucker_w),
            sav_all_tot_tot_w = sd(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat,names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat, names_prefix = 'sd_treat_', 
              values_from = value)

# table of t stats and p values for diff in means between T & C
complete_stats <- df_complete %>%
  select(treat, female_w, d_married_w, 
         worker_age_w, plucker_w, sav_all_tot_tot_w) %>%
  pivot_longer(cols=!treat,
               names_to='name',values_to='value') %>%
  group_by(name) %>%
  summarise(tstat = t.test(value ~ treat)$statistic,
            pval = t.test(value ~ treat)$p.value,
            .groups='keep')

# creating balance table from above tables
complete_bal <- complete_means %>% 
  left_join(complete_sds,
            by='name') %>%
  left_join(complete_stats,
            by='name')

# export as latex
complete_bal_tab <- xtable(complete_bal)
align(complete_bal_tab) <- xalign(complete_bal_tab)
display(complete_bal_tab) <- xdisplay(complete_bal_tab)
print(complete_bal_tab, include.rownames=FALSE)
```

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
name & treat\_0 & treat\_1 & sd\_treat\_0 & sd\_treat\_1 & tstat & pval \\ 
  \hline
female\_w & 0.37 & 0.33 & 0.48 & 0.47 & 1.21 & 0.23 \\ 
  d\_married\_w & 0.69 & 0.71 & 0.46 & 0.46 & -0.52 & 0.61 \\ 
  worker\_age\_w & 39.56 & 39.47 & 10.70 & 10.85 & 0.13 & 0.90 \\ 
  plucker\_w & 0.78 & 0.75 & 0.41 & 0.43 & 1.20 & 0.23 \\ 
  sav\_all\_tot\_tot\_w & 31968.41 & 32160.47 & 52691.10 & 49432.78 & -0.06 & 0.96 \\ 
   \hline
\end{tabular}
\end{table}


### 1.4

```{r}
# stratified randomization

# clear environment
rm(list = ls())

# set seed
set.seed(2090820)

# import data
df_strat <- read.csv('data/brune_et_al_data.csv')

n <- df_strat %>% nrow()
df_strat$rand <- runif(n, 0, 1)

# within each division, assign worker to treatment if random draw from 
# uniform[0,1] is greater than the bottom third of the draws of that division
# to produce a 2:1 T/C allocation
df_strat <- df_strat %>%
  group_by(division) %>%
  mutate(treat_strat = 
           ifelse(rand <= quantile(rand,c(0.33)), 0, 1)) %>% 
  ungroup()

df_strat %>% 
  count(division, treat_strat) %>% 
  pivot_wider(names_from = treat_strat, 
              names_prefix = 'treat_',
              values_from = n)
```

### 1.5

```{r results=F}
# table of means with each row as a variable and T/C as columns
strat_means <- df_strat %>%
  group_by(treat_strat) %>%
  summarise(female_w = mean(female_w),
            d_married_w = mean(d_married_w),
            worker_age_w = mean(worker_age_w),
            plucker_w = mean(plucker_w),
            sav_all_tot_tot_w = mean(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat_strat,names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat_strat, names_prefix = 'treat_', 
              values_from = value)

# table of standard deviations 
strat_sds <- df_strat %>%
  group_by(treat_strat) %>%
  summarise(female_w = sd(female_w),
            d_married_w = sd(d_married_w),
            worker_age_w = sd(worker_age_w),
            plucker_w = sd(plucker_w),
            sav_all_tot_tot_w = sd(sav_all_tot_tot_w),
            .groups = 'keep') %>%
  pivot_longer(cols=!treat_strat, names_to = 'name', values_to = 'value') %>%
  mutate(value = round(value,2)) %>%
  pivot_wider(names_from = treat_strat, names_prefix = 'sd_treat_', 
              values_from = value)

# table of t stats and p values for diff in means between T & C
strat_stats <- df_strat %>%
  select(treat_strat, female_w, d_married_w, 
         worker_age_w, plucker_w, sav_all_tot_tot_w) %>%
  pivot_longer(cols=!treat_strat,
               names_to='name',values_to='value') %>%
  group_by(name) %>%
  summarise(tstat = t.test(value ~ treat_strat)$statistic,
            pval = t.test(value ~ treat_strat)$p.value,
            .groups='keep')

# creating balance table from above tables
strat_bal <- strat_means %>% 
  left_join(strat_sds,
            by='name') %>%
  left_join(strat_stats,
            by='name') 

# export as latex
strat_bal_tab <- xtable(strat_bal)
align(strat_bal_tab) <- xalign(strat_bal_tab)
display(strat_bal_tab) <- xdisplay(strat_bal_tab)
print(strat_bal_tab, include.rownames=FALSE)
```

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
name & treat\_0 & treat\_1 & sd\_treat\_0 & sd\_treat\_1 & tstat & pval \\ 
  \hline
female\_w & 0.33 & 0.35 & 0.47 & 0.48 & -0.66 & 0.51 \\ 
  d\_married\_w & 0.70 & 0.70 & 0.46 & 0.46 & 0.16 & 0.88 \\ 
  worker\_age\_w & 39.29 & 39.63 & 11.02 & 10.64 & -0.43 & 0.67 \\ 
  plucker\_w & 0.74 & 0.78 & 0.44 & 0.42 & -1.06 & 0.29 \\ 
  sav\_all\_tot\_tot\_w & 33270.55 & 31461.38 & 48699.33 & 52228.98 & 0.50 & 0.61 \\ 
   \hline
\end{tabular}
\end{table}

## Survey CTO

### 2.1 

See pdf appended to the end of the doc.

### 2.2

```{r}
rand_treat_ids <- df_strat %>%
  filter(treat_strat == 1) %>%
  sample_n(10) %>% 
  select(worker_id, treat_strat)
rand_control_ids <- df_strat %>%
  filter(treat_strat == 0) %>%
  sample_n(10) %>%
  select(worker_id, treat_strat)
rand_ids <- bind_rows(rand_treat_ids,
          rand_control_ids)
rand_ids
survey_info_df <- df_strat %>%
       filter(worker_id %in% c(rand_treat_ids$worker_id,
                               rand_control_ids$worker_id)
              ) %>%
       select(worker_id, worker_age_w, follow_up_emp, follow_up_emp_Lujeri,
              follow_up_emp_other, follow_up_days_worked_Luj,
              follow_up_days_plucking, income_total_w,
              follow_up_large_purchases,
              FU_purch_varlist, FU_purch_livestock, FU_purch_metal_roof,
              FU_purch_home_imprv, FU_purch_phys_asset, FU_purch_new_house,
              follow_up_purchase_other)
```

Below I provide code which processes the data from my test survey (I manually inputted the data of the above workers).

```{r}
survey_df <- read.csv('data/survey_results.csv')

# share employed + avg inc for treatment and control
survey_df %>%
  left_join(rand_ids,
            by = 'worker_id') %>%
  group_by(treat_strat) %>%
  summarise(share_employed = mean(is_employed),
            avg_inc = mean(inc_last14),
            .groups = 'keep') %>%
  pivot_longer(cols = !treat_strat, names_to = 'name',
               values_to = 'value') %>%
  pivot_wider(names_from = treat_strat, names_prefix = 'treat_',
              values_from = value)
```



